---
title: "runcharter"
author: "John MacKintosh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{runcharter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The rationale for the package is to analyse multiple runcharts at once.
More importantly, the package will identify a run of n successive points on the desired side of the median line, and will recalculate / rebase the median accordingly.
Each time a run occurs, the median will be rebased.
This helps to track improvements over time. 


The default is to plot a faceted plot, allowing ggplot2 to handle the axis limits. 
You can specify the number of columns in the faceted plot, or let ggplot2 do its thing.



```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 5
)
```

Here are the function arguments and defaults. 
A default run of 9 is specified as that is the length of run used by the IHI for it's patient safety initiatives worldwide. 


```{r, eval=FALSE}
#function arguments and defaults

library(runcharter)
runcharter(df = signals,
          med_rows = 13,
          runlength = 9, 
          chart_title = NULL,
          chart_subtitle = NULL, 
          direction = "below",
          faceted = TRUE,
          facet_cols = NULL)

```




## Faceted plots - 1 column

```{r}
library(runcharter)

runcharter(signals,
                med_rows = 13,
                runlength = 9,
                chart_title = "facet test 1",
                chart_subtitle = " 1 column trellis using facet_wrap", 
                direction = "below",
                faceted = TRUE,
                facet_cols = 1)



```



```{r}

library(runcharter)

runcharter(signals,
                med_rows = 13,
                runlength = 9,
                chart_title = "facet test 2",
                chart_subtitle = " 2 column trellis using facet_wrap", 
                direction = "below",
                faceted = TRUE,
                facet_cols = 2)

```

## Design decisions

This package makes no attempt to identify or highlight 'trends' (consecutively increasing or decreasing points) or shorter 'shifts' of runs below the median.
In 8 years of working with QI data, I very rarely see trends, and when they do appear, they tend to result after a period of process degradation, when the process begins to return back to its natural median state. 

Research by qicharts2 author Jacob Anheoj shows that the trend rule is not reliable.

The package currently only identifies a run in one direction at a time. 
In other words, if you are trying to reduce adverse events, improvement will be evidenced with a run below the median. Therefore, you would not rebase the median if a run of points above the median was observed. 

If you want to rebase in both directions and / or  also apply the rules for trend then you should consider using Joseph Adam's 'runchart' package.

## We're not supposed to do this

"But [insert QI expert name here] says you should not automatically rebase the median unless you know the signal is genuine".

Yes, and no. 

Truth is, sometimes, there is too much information being collected, or the task of manually performing this analysis is very time consuming, especially if you have to make manual adjustments to spreadsheets / databases / graphical outputs, to deadlines, whilst battling traditional BI tools that were only ever designed for bar and pie charts.
Ultimately, I built this for me, to make my job easier. 
I believe other healthcare analysts, who work with QI data, will also find it beneficial.


## I don't like the default chart colours or axis limits

Use the package to do the analysis, take the outputs, and build your own plots.
Or save the output to a variable, retrieve the runchart component, and pass further ggplot2 syntax to amend the limits, labels etc.






